---
title: "Final"
author: "Anna MacFarlane, Edgar Salas, Jerry Fu, Taalin RaoShah"
date: "11/4/2020"
output:
  pdf_document: default
  html_document: default
--- 

```{r setup-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(plyr) #package for join command
library(maps) #package for world map
library(broom)
```

```{r load-datasets, message = FALSE, warning = FALSE}
report_2018 <- read_csv("data/2018.csv") %>% 
  mutate(year = "2018",
         `Perceptions of corruption` = as.numeric(`Perceptions of corruption`))
report_2019 <- read_csv("data/2019.csv") %>% 
  mutate(year = "2019")
country_region <- read_csv("data/2020.csv") %>% 
  mutate(year = "2020") %>% 
  select(`Country name`, `Regional indicator`)

names(country_region) <- c("Country", "Region")
names(report_2018) <- str_replace_all(names(report_2018), c(" " = "_"))
names(report_2019) <- str_replace_all(names(report_2018), c(" " = "_"))

worldhappiness <- full_join(report_2018, report_2019) 
worldhappiness <- worldhappiness %>% 
  rename(c("Country_or_region" = "Country")) %>% 
  left_join(country_region)

tidy_frame <- data.frame(matrix(unlist(worldhappiness), nrow = length(worldhappiness), byrow = T))

glimpse(worldhappiness)
```

```{r add-region}
#add regions to rows with missing Region column
worldhappiness <- worldhappiness %>% 
  mutate(Region = ifelse(Country == "Taiwan", "East Asia", Region),
         Region = ifelse(Country == "Qatar", "Middle East and North Africa", Region),
         Region = ifelse(Country == "Trinidad & Tobago", "Latin America and Caribbean", Region),
         Region = ifelse(Country == "Belize", "Latin America and Caribbean", Region),
         Region = ifelse(Country == "Northern Cyprus", "Middle East and North Africa", Region),
         Region = ifelse(Country == "Hong Kong", "East Asia", Region),
         Region = ifelse(Country == "Bhutan", "South Asia", Region),
         Region = ifelse(Country == "Somalia", "Middle East and North Africa", Region),
         Region = ifelse(Country == "Sudan", "Middle East and North Africa", Region),
         Region = ifelse(Country == "Angola", "Middle East and North Africa", Region),
         Region = ifelse(Country == "Syria", "Middle East and North Africa", Region),
         Region = ifelse(Country == "North Macedonia", "Central and Eastern Europe", Region))

worldhappiness
```

### Is there a relationship between region of the world and happiness score?

```{r world-map, fig.width = 10, fig.height = 5}
world_map <- map_data("world") %>% 
  mutate(region = ifelse(region == "USA", "United States", region)) %>% 
  mutate(region = ifelse(region == "Democratic Republic of the Congo", "Congo (Kinshasa)", region))
  
happiness_score_map <- left_join(worldhappiness, world_map, by = c("Country" = "region"))
ggplot(happiness_score_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Score), color = "white") +
  labs(title = "World Map", subtitle = "colored by Happiness Score", x = "", y = "") +
  scale_fill_viridis_c(option = "D")
```
https://www.datanovia.com/en/blog/how-to-create-a-map-using-ggplot2/

$H_0$: There is no association between region of the world and happiness score.
$H_a$: There is an association between region of the world and happiness score.
```{r chi-sq-region-score}
chisq.test(worldhappiness$Score, worldhappiness$Region, simulate.p.value = TRUE)
```
With a simulated p-value of 0.0004998 which is less than our $\alpha = 0.05$, we have
sufficient evidence to reject the null hypothesis in favor of the alternative hypothesis
and conclude that there is an association between region of the world and happiness score.

### Summary Statistics
```{r summary-statistics}
WE <- worldhappiness %>% 
  filter(Region == "Western Europe")
NorthAmerica <- worldhappiness %>% 
  filter(Region == "North America and ANZ")

worldhappiness %>% 
  filter(Region == "Western Europe") %>% 
  summarize(WE_mean_happiness = mean(Score), median_score = median(Score), SD = sd(Score))

worldhappiness %>%
  filter(Region == "North America and ANZ") %>%
  summarize(NA_mean_happiness = mean(Score), median_score = median(Score), SD = sd(Score))
```


###Create Function
```{r create-function}
diff_means <- function(variable, n_sims){
  boot_diffs <- numeric(n_sims)

  for(i in 1:n_sims){
    indices_WE <- sample(1:nrow(WE), replace = T)
    indices_NA <- sample(1:nrow(NorthAmerica), replace = T)
  
    temp_WE <- WE %>% 
      slice(indices_WE) %>% 
      summarize(mean_WE = mean(get(variable))) %>% 
      select(mean_WE) %>% 
      pull()
    temp_NA <- NorthAmerica %>% 
      slice(indices_NA) %>% 
      summarize(mean_NA = mean(get(variable))) %>% 
      select(mean_NA) %>% 
      pull()

    boot_diffs[i] <- temp_WE - temp_NA
  } 
  boot_diffs <- tibble(diffs = boot_diffs)

  offset <- boot_diffs %>% 
    summarize(offset = 0 - mean(diffs)) %>% 
    pull()
  null_dist <- boot_diffs %>% 
    mutate(centered_diffs = diffs + offset) %>% 
    select(centered_diffs)
  obs_diff <- boot_diffs %>% 
    summarize(obs_diff = mean(diffs)) %>% 
    pull()
  p_val <- null_dist %>% 
    mutate(extreme = ifelse(centered_diffs > abs(obs_diff), 1, 0)) %>% 
    summarize(p_val = mean(extreme)) 

  return(list(p_val = p_val, null_dist = null_dist, obs_diff = obs_diff))
}
```

### GDP per capita

```{r gdp}
set.seed(714)
worldhappiness %>% 
  filter(Region == "Western Europe") %>%
  summarize(WE_GDP = mean(GDP_per_capita))

worldhappiness %>%
  filter(Region == "North America and ANZ") %>%
  summarize(NorthA_GDP = mean(GDP_per_capita))

bootstrap <- diff_means(variable = "GDP_per_capita", 1000)

bootstrap$p_val
ggplot(bootstrap$null_dist, aes(x = centered_diffs)) +
  geom_histogram(binwidth = 0.005,
                 fill = "skyblue", 
                 color = "darkblue") +
  geom_vline(xintercept = bootstrap$obs_diff, color = "tomato", lwd = 2) +
  labs(title = "Simulated Null Distribution of Difference in GDP per capita", 
       x = "Difference in Mean",
       y = "Count")
```

### Social Support

```{r social-support}
set.seed(949)
worldhappiness %>% 
  filter(Region == "Western Europe") %>%
  summarize(WE_GDP = mean(Social_support))

worldhappiness %>%
  filter(Region == "North America and ANZ") %>%
  summarize(NorthA_GDP = mean(Social_support))

bootstrap <- diff_means(variable = "Social_support", 1000)

bootstrap$p_val

ggplot(bootstrap$null_dist, aes(x = centered_diffs)) +
  geom_histogram(binwidth = 0.005,
                 fill = "skyblue", 
                 color = "darkblue") +
  geom_vline(xintercept = bootstrap$obs_diff, color = "tomato", lwd = 2) +
  labs(title = "Simulated Null Distribution of Difference in Social Support", 
       x = "Difference in Mean",
       y = "Count")
```

### Freedom to make life choices 

$H_0$: $\mu_N$ <= $\mu_O$ 

$H_A$: $\mu_N$ > $\mu_O$ 

Null hypothesis: The mean freedom to make life choices score in Western Europe is less than or equal to the mean freedom to make life choices score in North America, Australia, and New Zealand.

Alternative hypothesis: The mean freedom to make life choices score in Western Europe is greater than the mean freedom to make life choices score in North America, Australia, and New Zealand.

```{r freedom-life-choices}
set.seed(1)

bootstrap <- diff_means(variable = "Freedom_to_make_life_choices", 1000)

bootstrap$p_val

ggplot(bootstrap$null_dist, aes(x = centered_diffs)) +
  geom_histogram(binwidth = 0.01,
                 fill = "skyblue", 
                 color = "darkblue") +
  geom_vline(xintercept = bootstrap$obs_diff, color = "tomato", lwd = 2) +
  labs(title = "Simulated null distribution of difference in mean 
freedom to make life choices scores",
       x = "Difference in mean", y = "Count")
```

### Generosity

$H_0$: $\mu_N$ <= $\mu_O$ 

$H_A$: $\mu_N$ > $\mu_O$ 

Null hypothesis: The mean generosity score in Western Europe is less than or equal to the mean generosity score in North America, Australia, and New Zealand.

Alternative hypothesis: The mean freedom to make life choices score in Western Europe is greater than the mean generosity score in North America, Australia, and New Zealand.

```{r generosity}
set.seed(2)

bootstrap <- diff_means(variable = "Generosity", 1000)

bootstrap$p_val

ggplot(bootstrap$null_dist, aes(x = centered_diffs)) +
  geom_histogram(binwidth = 0.01,
                 fill = "skyblue", 
                 color = "darkblue") +
  geom_vline(xintercept = bootstrap$obs_diff, color = "tomato", lwd = 2) +
  labs(title = "Simulated null distribution of difference in mean 
freedom to make life choices scores",
       x = "Difference in mean", y = "Count")
```

### Life Expectancy

```{r LE}
worldhappiness %>% 
  filter(Region == "Western Europe") %>%
  summarize(WE_life_expectancy = mean(Healthy_life_expectancy))

worldhappiness %>%
  filter(Region == "Western Europe") %>%
  summarize(RestofW_life_expectancy = mean(Healthy_life_expectancy))

bootstrap <- diff_means(variable = "Healthy_life_expectancy", 1000)

bootstrap$p_val

ggplot(bootstrap$null_dist, aes(x = centered_diffs)) +
  geom_histogram(binwidth = 0.01,
                 fill = "skyblue", 
                 color = "darkblue") +
  geom_vline(xintercept = bootstrap$obs_diff, color = "tomato", lwd = 2) +
  labs(title = "Simulated Null Distribution of Difference in 
  Mean Life Expectancy", 
       x = "Difference in Mean",
       y = "Count")
```

