---
title: "PerceptionsOfCorruption"
author: "Anna MacFarlane"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(plyr) #package for join command
library(broom)

knitr::opts_chunk$set(echo = TRUE)
report_2020 <- read_csv("data/2020.csv")
report_2018 <- read_csv("data/2018.csv") %>% 
  mutate(year = "2018") %>% 
  mutate(`Perceptions of corruption` = as.numeric(`Perceptions of corruption`))
report_2019 <- read_csv("data/2019.csv") %>% 
  mutate(year = "2019")

worldhappiness <- full_join(report_2018, report_2019)
#wrldhappy <- cbind(worldhappiness, report_2020$`Regional indicator`)
```

### Organizing data
```{r data}
corruption <- worldhappiness %>% 
  filter(! is.na(`Perceptions of corruption`)) %>% 
  select(Score, `Perceptions of corruption`, year)

year2019 <- corruption %>% 
  filter(year == "2019")
year2018 <- corruption %>% 
  filter(year == "2018")
```

### Correlational stats

```{r correlations}
corruption %>% 
  summarize(r_overall = cor(Score, `Perceptions of corruption`))

year2019 %>% 
  summarize(r_2019 = cor(Score, `Perceptions of corruption`))

year2018 %>% 
  summarize(r_2018 = cor(Score, `Perceptions of corruption`))

```

### Conditions for linear model 

```{r lm-conditions}
m_corrupt_aug <- augment(m_trees)
m_corrupt_aug %>%
  slice(1:3)
```
Independence
Question of whether countries are independent from each other...

Equal Variance, Linearity 
```{r equal-variance}
ggplot(m_corrupt_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(x = "Predicted overall score", y = "Residuals")
```

The spread of residuals does not appear to be equal around the line y=0, indicating a lack of equal variance and linearity.

Normality
```{r normality}
ggplot(m_corrupt_aug, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(x = "Residuals")

ggplot(m_corrupt_aug, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()
```
Both visualizations appear to be relatively normal.


Considering that the conditions for equal variance and linearity are clearly not met, while it is unclear whether the conditions for independence and normality are met, linear regression is not appropriate for this untransformed data. 

### Linear model and 95% confidence interval 

```{r linear-model-corrupt, message = FALSE}
# Linear model for relationship between perceptions of corruption score and overall happiness score
m_corrupt <- lm(Score ~ `Perceptions of corruption`, data = worldhappiness)

m_corrupt %>%
  tidy() %>%
  select(term, estimate) 


# Visualization of linear model
ggplot(data = corruption, aes(x = `Perceptions of corruption`,
                              y = Score)) + 
  geom_point(size = 0.8) + 
  geom_smooth(method = lm, color = "blue", size = 0.5) +
  labs(title = "Linear model for relationship between perceptions of corruption
and overall happiness score",
       subtitle = "Moderate, positive assocation",
       x = "Perceptions of corruption score",
       y = "Overall happiness score")

#Confidence interval for linear model above 
broom::confint_tidy(m_corrupt, conf.level = 0.95) %>% 
  slice(2)

```
Overall happiness score = 4.873 + 4.616(Perceptions of corruption)

The confidence interval generated is about (3.416, 5.817). We are 95% confident that for every additional point increase in the perceptions of corruption, the overall happiness score with increase, on average, by 4.616 points. 


### Hypothesis test for association

** Note: Is this dumb to do because obviously they're associated because they used it to calculate the overall happiness score?

This test will be evaluated at the $\alpha = 0.05$ level.

Null hypothesis: There is no association between the perceptions of corruption and the overall happiness score. They are independent.

Alternative hypothesis: There is an association between the perceptions of corruption and the overall happiness score. They are not independent.


```{r association}
tidy(m_corrupt)
```
Considering that our p-value is less that $\alpha =$ 0.05, we can reject the null hypothesis. There is statistically significant evidence to suggest that there is an assocation between perceptions of corruption and the overall happiness score. They are not independent. 


### Multiple regression for all the factors we are evaluating

```{r multiple-reg-for-all}
m_happy <- lm(Score ~ `Perceptions of corruption` + `Social support` +
                `GDP per capita`, data = worldhappiness)

m_happy %>% 
  tidy() %>%
  select(term, estimate) 
  
```
